---
title: "Exercise_2"
output: html_document
---
```{r}
##-----------------------------------------------##
##    Author: Adela Sobotkova                    ##
##    Institute of Culture and Society           ##
##    Aarhus University, Aarhus, Denmark         ##
##    adela@cas.au.dk                             ##
##-----------------------------------------------##

#### Goals ####

# - Understand the provided datasets
# - Learn how to reproject spatial data
# - Limit your data into an area of interest
# - Create a new map

# We highlighted all parts of the R script in which you are supposed to add your
# own code with: 

# /Start Code/ #

print("Hello World") # This would be your code contribution

# /End Code/ #

#### Required R libraries ####

# We will use the sf, raster, and tmap packages.
# Additionally, we will use the spData and spDataLarge packages that provide new datasets. 
# These packages have been preloaded to the worker2 workspace.

library(sf)
library(raster)
library(tmap)
library(spData)
library(spDataLarge)

#### Data sets #### 

# We will use two data sets: `srtm` and `zion`.
# The first one is an elevation raster object for the Zion National Park area, and the second one is an sf object with polygons representing borders of the Zion National Park.

srtm <- raster(system.file("raster/srtm.tif", package = "spDataLarge"))
zion <- read_sf(system.file("vector/zion.gpkg", package = "spDataLarge"))

# Additionally, the last exercise (IV) will used the masked version of the `lc_data` dataset.

study_area <- read_sf("data/study_area.gpkg")
lc_data <- raster("data/example_landscape.tif")
lc_data_masked <- mask(crop(lc_data, study_area), study_area)

```
## Exercise 1

```{r}

# 1. Display the `zion` object and view its structure.
# What can you say about the content of this file?
# What type of data does it store? 
# What is the coordinate system used?
# How many attributes does it contain?
# What is its geometry?

str(zion)
```
### Question 1
Using the "View()" function, we see that the file only contains one row, one entry if you will. The biggest field belongs to the 'geom' column in which we get to see the data for the entire polygon that outlines Zion National Park. The file also contains meta-data, where one can see the state, region, and who created it amongst other things. 

Expanding the last column (geom), we can see that the polygon is represented in an 2D XY-coordinate system which has 11 attributes.
```{r}
# 2. Display the `srtm` object and view its structure.
# What can you say about the content of this file? 
# What type of data does it store?
# What is the coordinate system used? 
# How many attributes does it contain?
# How many dimensions does it have? 
# What is the data resolution?

# Your solution (type answer to the questions as code comments and the code used)

res(srtm)
str(srtm)
srtm
```
### Question 2
This file is a 457x465 raster with a resolution of ~0.0008. It is a 3D CRS with 12 slots (attributes), each containing sub-attributes. 

## Exercise 2
```{r}
#### Exercise II ####

# 1. Reproject the `srtm` dataset into the coordinate reference system used in the `zion` object. 
# Create a new object `srtm2`
# Vizualize the results using the `plot()` function.
# 2. Reproject the `zion` dataset into the coordinate reference system used in the `srtm` object.
# Create a new object `zion2`
# Vizualize the results using the `plot()` function.



```


